Envelope Estimation using Geometric Properties of a Discrete Real Signal

### 2020-12-06 Papers/01_Pulses_Envelope/PulsesEnvelope.tex
revised

\begin{abstract}
% Background
Despite being an elusive concept, the temporal amplitude envelope of a signal is essential for its complete characterization, being the primary information-carrying medium in spoken voice and telecommunications, for example. Intuitively, the temporal envelope can be understood as a slowly varying function that multiplies the signal, being responsible for its outer shape. 
Envelope detection techniques have applications in areas like health, sound classification and synthesis, seismology and speech recognition. 
% Problem | Objectives
Nevertheless, a general approach to digital envelope detection of signals with rich spectral content doesn't exist, as most methods involve manual intervention, in the form of filter design, smoothing, as well as other specific design choices, based on prior knowledge of the nature of the specific waves under investigation.
To address this problem, we propose a framework that uses intrinsic characteristics of a signal to estimate its envelope, eliminating the necessity of parameter tuning.
% Methods | Procedure | Approach
The approach here described draws inspiration from geometric concepts to estimate the temporal envelope of an arbitrary signal; specifically, a new measure of discrete curvature is used to obtain the average radius of curvature of a discrete wave, that will as a threshold to identify the wave's samples that are part of the envelope, in a manner similar to the alpha shapes approach to the identification of the concave hull of a set of points.
% Results | Findings | Product
The algorithm compares favourably with classic envelope detection techniques based on smoothing, filtering and the Hilbert Transform, besides being physically plausible.
We provide visualizations of the envelope extracted via the algorithm for various real-world signals, with very diverse characteristics, such as voice, spoken and sang, and pitched and non-pitched musical instruments, and discuss some approaches to assess the quality of the obtained envelopes.
A Python module, installable via the Python Package Index (PyPI), with a pure Python and a specialized C++ implementation for 64-bit Windows systems, interactive visualizations of envelopes for a diverse range of digital waves, as well as the source code for the Python implementation, are available online at \textcite{2020TarjanoEnvelope}.
% Conclusions | Implications
Besides the most direct applications of this work to audio classification and synthesis, we foresee impact in compression techniques and machine learning approaches to audio. The discrete curvature definition presented could also be extended to three-dimensional settings, to improve shape detection algorithms based on alpha shapes.
\end{abstract}

@@

In many contexts, however, the temporal amplitude envelope of a signal plays a prominent role: according to \textcite{2017QiRelative}, for example, the envelope is at least as important as the fine structure of a sound wave in the context of the intelligibility of Mandarin tones. This is also the case for the English language \parencite{1995ShannonSpeech}, where even envelopes modulating mostly noise were still capable of conveying meaning.

The envelope also helps to impart emotion and identity to the human voice \parencite{2018ZhuContributions}, and envelope preserving characteristics in concert halls are associated with their pleasantness \parencite{2011LokkiEngaging}.

@@

One direct way of defining the shape of a set of points in $ \mathbb{R}^2 $ is using the convex hull: as defined in \textcite{2008BergComputational}, the convex hull of a set of points $ S $ in two-dimensional space is the smallest convex subset of the plane that contains all the points in $ S $. Convex hulls are unique, in the sense that a set of points define one and only one convex hull.

Generalizing that definition, \textcite{1983Edelsbrunnershape} introduced the concept of alpha shapes, that can be seen as a Concave Hull; a mathematically well-defined generalization of the Convex Hull of a finite set of points, closely related to the Delaunay triangulation and Voronoi diagrams of those points.

The intuition behind the alpha shapes algorithm is that only points that can be touched by a circle of radius $ \alpha $ coming from an infinite distance towards the set of points of interest are part of the frontier, that is, of the Concave Hull of this set.

Once an $ \alpha $ parameter is defined alpha shapes conserve the uniqueness property, besides reducing to the Convex Hull when $ \alpha = \infty $. Consider, for example, the artificially generated discrete wave in figure \ref{fig:Hulls}
below, where the links between consecutive samples were deemphasized to highlight the “set of points” nature of a signal: The Convex Hull and a Concave Hull, in the form of an alpha shape with a particular value of $ \alpha $ are illustrated.

\begin{figure}[ht!]
  \centering
    \includegraphics{images/00ConvexHull.pdf}
  \caption{A discrete signal interpreted as a set of points in the Cartesian plane, and the Convex Hull and Concave Hull at $ \alpha=300 $ defined by it.}
  \label{fig:Hulls}
\end{figure}

The alpha shapes approach is used in areas such as the detection of features in images \parencite{2016VarytimidisAlpha}, reconstruction of surfaces from a cloud of points \parencite{2015WuAutomated} and Spectroscopy \parencite{2019XuModeling}, with the last work, which involves the estimation and removal of the Blaze function (a kind of envelope) of an echelle spectrograph, being particularly illustrative of the potential of synergy between geometric and DSP approaches.

Other steps in the direction of translating geometric algorithms to the context of envelope detection were also made by \textcite{2015YangSkeleton} via an algorithm based on the construction of a skeleton underlying the digital wave of interest, and also via the direct translation of computer vision methods to the task of envelope detection \parencite{2015YangRepresenting}.

@@

The rest of this paper is divided as follows: 
We start by trying to define as specifically as possible what a temporal envelope is, especially in the context of this work, providing a mathematical definition, in terms of the Hilbert transform, for the case of a narrowband signal.

We proceed to present the representation of a discrete wave used in this work, largely based on the concept of dividing the wave into its constituent pulses, and how it can simplify the algorithm, reducing the high dimensionality generally present by digital representations of waves.

Next, we review methods for the estimation of the curvature of a discrete wave, after which we present our own; this method will then be used to estimate the equivalent circle, that will be used to define the envelope of the wave.

In the results section, we present some illustrations of the envelopes obtained with our method, comparing both accuracy and running time with the most common algorithms encountered in the literature. We also suggest an alternative use of the algorithm to identify the negative and positive frontiers of a wave, a kind of partial envelope considering only the positive and negative pulses of a wave, respectively.

We follow with some suggestions about how the approach here presented can be useful beyond envelope detection, exemplifying with some applications, such as the extractions of the average waveform of a digital wave, concluding with some commentary on potential future developments of the research here presented.

@@

The interest in the theory of envelope detection arose in the analogue domain, with the widespread adoption of radio communications, where the problem is typically restricted to simple sinusoids, often with well-defined frequencies \parencite{2011TurnerDemodulation}.

In this context, the most mathematically sound definition of an envelope involves the representation of its underlying wave as an analytic signal, introduced by Gabor in 1946 \parencite{2007HahnHistory}. In his work, \textcite{1946GaborTheory} applies the then relatively new mathematical machinery of the quantum mechanics to unify the time and frequency-domain representations of a wave, showing how the Hilbert transform could be applied to a real signal in order to obtain an equivalent complex signal, later known as the analytic signal.

This analytic signal has the form $ A(t) = S(t) + \mathbb{H}(S(t)) \ \text{i} $ \parencite{2016HePraat} where $ S(t) $ is the original real signal. $ \mathbb{H}(S(t)) $, the Hilbert transform of the original signal, becomes the imaginary part of the analytic signal; the envelope of a signal thus represented can be straightforwardly obtained by the computation of the complex modulus of $ \mathbb{H}(S(t)) $.

@@

Figure \ref{fig:AnalyticSignal} exemplifies that, showing the envelope obtained via the analytic signal, for a pure sinusoid with a local frequency of 20 cycles, modulated by a polynomial of the third degree, illustrating the difference in the shape of the envelope in the presence and absence of Gaussian white noise with standard deviation of $ 1/10 $ of the wave's maximum amplitude.

It is readily noticeable from figure \ref{fig:AnalyticSignal} that, as soon as noise is introduced in the original signal, it is reflected in the envelope.

%%%%%%%%%%%%%%%%%

\begin{figure}[ht!]
  \centering
    \includegraphics{images/01AnalyticSignal.pdf}
  \caption{Envelope of a pure sinusoid modulated by a polynomial of degree 3, as obtained by the Hilbert Transform approach. In the first half, the sinusoid is free of noise, while in the second half Gaussian noise with a standard deviation of $ 1/10 $ of the maximum amplitude of the wave was added to the base signal.}
  \label{fig:AnalyticSignal}
\end{figure}

@@

Also, in this work we are concerned with the discrete version of this problem: given a finite digital wave represented by the vector $ \textbf{w} \in \mathbb{R}^n $, instead of the continuous function $ W(t) $, obtaining the temporal envelope of $ \textbf{w} $.

@@

\begin{align} \label{eq:Envelope}
  & \textbf{w}, \textbf{e}, \textbf{c} \in \mathbb{R}^n \\
  & \textbf{w} = (w_0, w_1, \cdots, w_{n-1}) \\
  & \textbf{e} = (e_0, e_1, \cdots, e_{n-1}) \\
  & \textbf{c} = (c_0, c_1, \cdots, c_{n-1}) \\
  & \textbf{w} = \textbf{e} \odot \textbf{c} \\
  & w_i = e_i c_i \quad \forall \quad i, 0 \le i \le n-1
\end{align}

The $ \odot $ operator in \ref{eq:Envelope} stands for the Hadamard product, denoting element wise multiplication of two vectors. Figure \ref{fig:Envelope} provides an example of the vectors just defined: \textbf{w} is obtained by modulating the normalized carrier wave \textbf{c} using the smooth envelope \textbf{e}.

Note from Figure \ref{fig:Envelope} that, at some local maxima, the envelope is defined, being in fact equal to the value of the wave \textbf{w}, since $ c_i = 1 $ at those points, rendering $ e_i = 1 w_i $. The carrier wave \textbf{c} is periodic, and normalized between $ \{-1, 1\}$. Each cycle of the wave, then, reaches the value 1 only at its maximum making $ w_i$ equal to the value of the envelope at that point.

Image \ref{fig:Envelope}, along with the Concave Hull illustrated in figure \ref{fig:Hulls}, helps forming the intuition behind the method developed in this work: the idea is to identify the local extrema that touch the envelope (or, to put in another way, the local extrema in relation to the whole wave that are global extrema inside their own pseudo-cycles), filtering out those who don't; in this way, we're defining the envelope itself at those points.

In this work, this is accomplished with the use of a circle of a defined radius, large enough to prevent it from reaching the smallest local extrema when "rolled" around the (geometric) representation of the wave: only the extrema touched by the circle will be marked as part of the envelope.

In the next section, such a geometric representation will be formulated, along with the necessary tool to estimate the average curvature of the wave and, consequently, the appropriate radius of the circle.

\begin{figure}[ht!]
  \centering
    \includegraphics{images/02Envelope.pdf}
  \caption{Example of a discrete wave \textbf{w} composed by the element wise multiplication of an envelope \textbf{e} and a carrier \textbf{c}. The local extrema of \textbf{w} are highlighted with a circle.}
  \label{fig:Envelope}
\end{figure}

@@

Instrumental to the method here presented is the definition of a pulse: each time the value of \textbf{w} changes sign, that is, every time the discrete wave \textbf{w} represents crosses the horizontal axis, the beginning of a pulse is defined, with the next crossing defining its end. This definition is in line with the one presented in \textcite{national1996telecommunications}, where a pulse is defined as a rapid change in the amplitude of a signal, followed by a fast return to the baseline value; zero, in our case.

From \ref{fig:Envelope} we can see that only the maximum point of a positive pulse or the minimum point of a negative one can potentially be equal to the envelope and are, thus, our only points of interest. We can then proceed with the rest of the method considering only those points, to great computational advantage.

@@

This relation is illustrated in figure \ref{fig:Pulses}. We call it P to emphasize the connection with the pulses of \textbf{w}, noting that P is, in fact, a set of points.

@@

P is not fit for geometric interpretation yet, since the abscissa and ordinate of his orthogonal coordinate system have different units: in the vertical axis, we have a unit related to the instantaneous intensity of the wave, while in the horizontal axis we have the index at which each extremum occurs, ultimately a time-related unit.

@@

In this new coordinate system we have a basis vector $ \mathbf{b}^\prime = \{ \mathbf{b}^\prime_x, \mathbf{b}^\prime_y \}$ where $ \mathbf{b}^\prime_x = (i, 0) $ and $\mathbf{b}^\prime_y = (0, i) $, related to the original basis vector by:

@@

The effect of this normalization can be seen in the points shown in figure \ref{fig:Vectors}, shown in scale in both coordinate systems. The set V of the vectors from each point of P to the next is also shown in the picture, as it will be used to estimate the average curvature of P, we define this set below.

@@

Concerning direct methods \textcite{2014CarrollSurvey}, for example, derive three such definitions based on the approximation of a circle by an inscribed, centered and circumscribed polygon. In the context of tridimensional meshes, \textcite{2016VasaMultivariate} evaluate a range of existing estimators from a multivariate point of view.

Those approaches define the discrete curvature in relation to the vertices of a wave. In our particular case, it wouldn't make sense to talk about the curvature of a single pulse, since it is what our points ultimately represent. Instead, we are interested in the change of direction, what the term curvature ultimately means, between two adjacent pulses, as the vectors in the set V suggests.

@@

The rationale is to find, for each vector $ v_k \in \text{V} $, the radius $ r_k $ of the equivalent circle whose tangent has the same change in direction, in the same horizontal distance, as the vector of interest, as shown in figure \ref{fig:DiscreteCurvature}. The average curvature of the P will then be obtained by the average of all radii $ r_k $.

\begin{figure}[ht!]
  \centering
    \includegraphics{images/05DiscreteCurvature.pdf}
  \caption{The tangent unit vector of the circle changes from the horizontal direction in $ \hat{u}_0 $ to an inclination of $ \theta $ in $ \hat{u}_1 $, $ \theta $ being the angle that the vector $ v_k $ makes with the horizontal direction.}
  \label{fig:DiscreteCurvature}
\end{figure}

From figure \ref{fig:DiscreteCurvature} is easy to see that $ r_k = v_{k,x} \sin(\theta_k)$ where $ v_{k,x} $ is the component of $ v_k $ in the horizontal direction; $ \theta_k $ itself can be obtained using the slope $ m_0 = 0 $ of any line in the horizontal direction and the slope $ m_k = v_{k,y} / v_{k,x} $ defined by $ v_k $ via the equality $ \theta_j = \arctan \big( (m_k - m_0) / (1 + m_k m_0) \big) $.

The radius of the circle that represents the average curvature of V, and will be used to obtain the envelope of the discrete wave \textbf{w} can be obtained via equation \ref{eq:Radius}:

@@

We need now to construct an algorithm, using the radius just obtained, to identify the points that belong to the envelope.
Generally, the algorithm for the alpha shapes approach resorts to the construction of the Delaunay triangulation of the set of points, that is latter filtered. We can adopt a more straightforward method, as our points have a defined structure:

@@

For all circles of center $ x_o, y_o $ and radius $ r $ that can be constructed to pass through any two different points of P, that is, $ \forall \quad P_b, P_c $ such that $ (P_{b,x} - x_o)^2 + (P_{b,y} - y_o)^2 = (P_{c,x} - x_o)^2 + (P_{c,y} - y_o)^2 = r^2 $, $ P_a $ will be part of the set $ E $ of the points belonging to the envelope if and only if none of those circles contain $ P_a $, or $ P_a \in \text{E} \leftrightarrow (P_{b,x} - x_o)^2 + (P_{b,y} - y_o)^2 > r^2 $. Formalizing, we have:

@@

The algorithm follows directly from the definition in \ref{eq:InEnvelope} after noting that the first and last members of P will always be part of the envelope, since they are reachable a circle of any radius $ r $. From $ P_0 $, our first pivot point, we construct circles passing through $ P_1, P_2, \cdots, P_d $ until a circle that doesn't contain any point in P is found; $ P_d $ then becomes the new pivot point and is included in E, and this procedure continues until $ P_{m-1} $ is reached.

The procedure is formalized in algorithm \ref{FindEnvelope}. We don't provide similar algorithmic descriptions of the preceding steps since the Python implementation can easily serve as pseudocode.

@@

The set E obtained via algorithm \ref{FindEnvelope} is a set of points. The most straightforward way to transform those points into the vector $ \textbf{e} \in \mathbb{R}^n $ is via linear interpolation, the approach used in this work. 
It is worth noting, however, that many other procedures are available, both for the interpolation and the smooth approximation of points. We deliberately choose linear interpolation to avoid blurring the results of the method.

\section{Results}
Given the lack of a precise definition of what an envelope should be, we begin this section suggesting an empirical metric, based on the behaviour of the carrier wave \textbf{c}. We also comment on some metrics presented in the literature, and compare our algorithm with traditional demodulation approaches, both in relation to execution time and accuracy; to this last end, a numerical indicator is introduced.

@@

There are no agreed-upon objective metrics to assess the quality of an envelope. Recalling figure \ref{fig:Envelope}, however, and the accompanying discussion, an intuitive way is to consider the carrier wave \textbf{c}, and how well it seems to be bounded in amplitude in the interval $ \{-1, 1\} $. To that end, we provide the following figures of original discrete waves \textbf{w}, their envelope \textbf{e}, and the inferred carrier, obtained after dividing, element-wise, the original wave by its envelope.

From figure \ref{fig:spoken_voice} it can be seen that the algorithm performs well under the carrier metric. This is especially remarkable for the case of spoken voice, a notorious complex sound with fast changes in pitch and high noise content.

These characteristics are in stark contrast with the sang voice in figure \ref{fig:singer_voice}, where a recording of an alto singer is shown; there, the wave is almost periodic, with a stable waveform.

@@

\begin{figure}[ht!]
  \centering
    \includegraphics{images/06Graph_spoken_voice.pdf}
  \caption{Signal, envelope and carrier for a record of male voice uttering the word "amazing".}
  \label{fig:spoken_voice}
\end{figure}

@@

The first one states that if a signal is bounded in magnitude, then its envelope should be, as well. Our algorithm complies with this requisite, as the envelope is composed of selected samples from the original signal itself.

@@

Nevertheless, if we picture both the wave and the envelope as samples from continuous functions, is easy to see that this t is the case for our algorithm; since the envelope is composed by some samples of the digital wave, in the limit where all the samples of the wave are also part of the envelope, we have that the maximum frequency for both is the same.

@@

The third condition states what, for a periodic signal, the envelope should be a straight line with the intercept equal to the amplitude of this signal. Since all the local maxima have the same amplitude, our envelope will indeed be a straight line, with the same amplitude as the original wave.

@@

Direct comparison with many of the more recent algorithms is made difficult by the unavailability of digital implementations of such works, many designed to process analogue signals \parencite[e.g.,][]{2018AssefModeling}.

>>>>

\begin{abstract}
% Background
Despite being an elusive concept, the temporal amplitude envelope of a signal is essential for its complete characterization, being the primary information-carrying medium in spoken voice and telecommunications, for example. Intuitively, the temporal envelope can be understood as a slowly varying function that multiplies the signal, being responsible for its outer shape. 
Envelope detection techniques have applications in areas like medicine, sound classification and synthesis, seismology and speech recognition.
% Problem | Objectives
Nevertheless, a general approach to digital envelope detection of signals with rich spectral content doesn't exist, as most methods involve manual intervention, in the form of filter design, smoothing, and other specific design choices, based on prior knowledge of the signals under investigation.
To address this problem, we propose an algorithm that uses intrinsic characteristics of a signal to estimate its envelope, eliminating the necessity of parameter tuning.
% Methods | Procedure | Approach
The approach here described draws inspiration from geometric concepts to estimate the temporal envelope of an arbitrary signal; specifically, a new measure of discrete curvature is used to obtain the average radius of curvature of a discrete wave, that will serve as a threshold to identify the wave's samples that are part of the envelope. 
% Results | Findings | Product
The algorithm compares favourably with classic envelope detection techniques based on smoothing, filtering and the Hilbert Transform, besides being physically plausible.
We provide visualizations of the envelope extracted via the algorithm for various real-world signals, with very diverse characteristics, such as voice, spoken and sang, and pitched and non-pitched musical instruments, and discuss some approaches to assess the quality of the obtained envelopes.
A Python module implementing the algorithm was made available via the Python Package Index (PyPI); interactive visualizations of envelopes for a diverse range of digital waves, as well as the source code for the Python implementation, are available online at \textcite{2020TarjanoEnvelope}.
% Conclusions | Implications
Besides the most direct applications of this work to audio classification and synthesis, we foresee impact in compression techniques and machine learning approaches to audio. We briefly discuss some potential paths in this direction. The discrete curvature definition presented could also be extended to three-dimensional settings, to improve shape detection algorithms based on alpha shapes.
\end{abstract}

@@

In many contexts, however, the temporal amplitude envelope of a signal plays a prominent role: according to \textcite{2017QiRelative}, for example, the envelope is at least as important as the fine structure of a sound wave in the context of the intelligibility of Mandarin tones. This is also the case for the English language \parencite{1995ShannonSpeech}, where even envelopes modulating mostly noise are still capable of conveying meaning.

The envelope also helps to impart emotion and identity to the human voice \parencite{2018ZhuContributions} and music: Concert halls with envelope preserving characteristics are deemed more pleasant \parencite{2011LokkiEngaging}, for example.

@@

One direct way of defining the shape of a set of points in $ \mathbb{R}^2 $ is using the concept of a convex hull: as defined in \textcite{2008BergComputational}, the convex hull of a set of points $ S $ in two-dimensional space is the smallest convex subset of the plane that contains all the points in $ S $. Convex hulls are unique, in the sense that a set of points define one and only one convex hull.

Generalizing that definition, \textcite{1983Edelsbrunnershape} introduced the concept of alpha shapes, that can be seen as a concave hull; a mathematically well-defined generalization of the convex hull of a finite set of points, closely related to the Delaunay triangulation and Voronoi diagrams of those points.

The intuition behind the alpha shapes algorithm is that only points that can be touched by a circle of radius $ \alpha $ coming from an infinite distance towards the set of points of interest are part of the frontier, that is, of the concave hull of this set.

Once a radius $ \alpha $ is defined, alpha shapes conserve the uniqueness property, in fact tending to the convex hull when $ \alpha \to \infty $. Consider, for example, the artificially generated discrete wave in figure \ref{fig:Hulls} below, where the links between consecutive samples were deemphasized to highlight the “set of points” nature of a signal: The convex hull and a concave hull, in the form of an alpha shape with a particular value of $ \alpha $ are illustrated.

\begin{figure}[ht!]
  \centering
    \includegraphics{images/00ConvexHull.pdf}
  \caption{A discrete signal interpreted as a set of points in the Cartesian plane, and the convex and concave hull at $ \alpha=300 $ defined by it.}
  \label{fig:Hulls}
\end{figure}

The alpha shapes approach is used in areas such as detection of features in images \parencite{2016VarytimidisAlpha}, reconstruction of surfaces from a cloud of points \parencite{2015WuAutomated} and spectroscopy \parencite{2019XuModeling}.

This last work, which involves the estimation and removal of the Blaze function (a kind of envelope) of an echelle spectrograph, being particularly illustrative of the potential synergy between geometric and DSP approaches.

Other steps in the direction of translating geometric algorithms to the context of envelope detection were made by \textcite{2015YangSkeleton} via an algorithm based on the construction of a skeleton underlying the digital wave of interest, and also via the direct translation of computer vision methods to the task of envelope detection \parencite{2015YangRepresenting}.

@@

The rest of this paper is divided as follows: We start by defining as specifically as possible what a temporal envelope is, especially in the context of this work and providing a mathematical definition, in terms of the Hilbert transform, for the case of a narrowband signal.

We proceed to illustrate the representation of a discrete wave used in this work, largely based on the concept of dividing the wave into its constituent pulses, and how it can simplify the algorithm, reducing the high dimensionality generally present in digital representations of waves.

Next, we review methods for the estimation of the curvature of a discrete wave, after which we present our own; this method will then be used to estimate the equivalent circle, that will be used to define the envelope of the wave, via the procedure described in sequence.

In the results section, we present some illustrations of the envelopes obtained with our method, comparing both accuracy and running time with the most common algorithms encountered in the literature. We also suggest an alternative use of the algorithm to identify the negative and positive frontiers of a wave, that can be seen as partial envelopes considering only the positive and negative pulses of a wave, respectively.

We follow with suggestions about how the approach here presented can be useful beyond envelope detection, exemplifying with some applications, such as the extractions of the average waveform of a digital wave, and concluding with commentary on potential future developments of the method.

@@

The interest in the theory of envelope detection arose in the analog domain, with the widespread adoption of radio communications, where the problem is typically restricted to simple sinusoids, often with well-defined frequencies \parencite{2011TurnerDemodulation}.

In this context, the most mathematically sound definition of an envelope involves the representation of its underlying wave as an analytic signal, as introduced by Gabor in 1946 \parencite{2007HahnHistory}. In his work, \textcite{1946GaborTheory} applies the then relatively new mathematical machinery of the quantum mechanics to unify the time and frequency-domain representations of a wave, showing how the Hilbert transform could be applied to a real signal in order to obtain an equivalent complex signal, later known as the analytic signal.
This analytic signal has the form $ A(t) = S(t) + \mathbb{H}(S(t)) \ \text{i} $ \parencite{2016HePraat} where $ S(t) $ is the original real signal. $ \mathbb{H}(S(t)) $, the Hilbert transform of the original signal, becomes the imaginary part of the analytic signal; the envelope of a signal thus represented can be straightforwardly obtained by the computation of the complex modulus of $ \mathbb{H}(S(t)) $.

@@

Figure \ref{fig:AnalyticSignal} exemplifies the Hilbert envelope, for a pure sinusoid modulated by a polynomial, illustrating the difference in the shape of the envelope in the presence and absence of Gaussian white noise. It is readily noticeable from the figure that, as soon as noise is introduced in the original signal, it is reflected in the envelope.

\begin{figure}[ht!]
  \centering
    \includegraphics{images/01AnalyticSignal.pdf}
  \caption{Envelope of a pure sinusoid with a local frequency of 20 cycles modulated by a polynomial of degree 3, as obtained by the Hilbert Transform approach. In the first half, the sinusoid is free of noise, while in the second half Gaussian noise with a standard deviation of $ 1/10 $ of the maximum amplitude of the wave was added to the base signal.}
  \label{fig:AnalyticSignal}
\end{figure}

@@

In this work we are concerned with the discrete version of this problem: given a finite digital wave represented by the vector $ \textbf{w} \in \mathbb{R}^n $, instead of the continuous function $ W(t) $, obtaining the temporal envelope of $ \textbf{w} $.

@@

\begin{align} \label{eq:Envelope}
  & \textbf{w}, \textbf{e}, \textbf{c} \in \mathbb{R}^n \\
  & \textbf{w} = (w_0, w_1, \cdots, w_{n-1}) \\
  & \textbf{e} = (e_0, e_1, \cdots, e_{n-1}) \\
  & \textbf{c} = (c_0, c_1, \cdots, c_{n-1}) \\
  & \textbf{w} = \textbf{e} \odot \textbf{c} \quad \therefore \quad w_i = e_i c_i \quad \forall \quad i, 0 \le i \le n-1
\end{align}

The $ \odot $ operator in \ref{eq:Envelope} stands for the Hadamard product, denoting element-wise multiplication of two vectors. Figure \ref{fig:Envelope} provides an example of the vectors just defined: \textbf{w} is obtained by modulating the normalized carrier wave \textbf{c} using the envelope \textbf{e}.

Note from Figure \ref{fig:Envelope} that, at some local maxima, the envelope is well-defined, being in fact equal to the value of the wave $ w_i $, since $ c_i = 1 $ at those points, rendering $ e_i = 1 w_i $.

The carrier wave \textbf{c} in the figure is periodic, and normalized between $ \{-1, 1\}$. Each cycle of the wave, then, reaches the value 1 only at its maximum making $ w_i $ equal to the value of the envelope at one point in each cycle.

Image \ref{fig:Envelope}, along with the concave hull illustrated in figure \ref{fig:Hulls}, helps forming the intuition behind the method developed in this work: the idea is to identify the local extrema that touch the envelope, those being also the extrema of each (pseudo-)cycle.

This is accomplished with the use of a circle with a defined radius, large enough to prevent it from reaching the smallest local extrema when "rolled" around the (geometric) representation of the wave. Only the extrema touched by the circle will be marked as part of the envelope.

In the next section, a compatible geometric representation will be formulated, along with the necessary tools to estimate the average curvature of the wave and, consequently, the appropriate radius of the circle.

\begin{figure}[ht!]
  \centering
    \includegraphics{images/02Envelope.pdf}
  \caption{Example of a discrete wave \textbf{w} composed by the element-wise multiplication of an envelope \textbf{e} and a carrier \textbf{c}. The local extrema of \textbf{w} are highlighted with a circle.}
  \label{fig:Envelope}
\end{figure}

@@

Instrumental to the method here presented is the definition of a pulse: each time the value of \textbf{w} changes sign, that is, every time the discrete wave \textbf{w} crosses the horizontal axis, the beginning of a pulse is defined, with the next crossing defining its end. This definition is in line with the one presented in \textcite{national1996telecommunications}, where a pulse is defined as a rapid change in the amplitude of a signal, followed by a fast return to the baseline value; zero, in our case.

From \ref{fig:Envelope} and the discussion in the preceding chapter, we can see that only the maximum point of a positive pulse or the minimum point of a negative one can potentially be equal to the envelope and are, thus, our only points of interest. We can then proceed, for the rest of the method, considering only those points, to great computational economy.

@@

This relation is illustrated in figure \ref{fig:Pulses}. We call it P to emphasize the connection with the pulses of \textbf{w}.

@@

The points in P are not fit for geometric interpretation yet, since the abscissa and ordinate of their orthogonal coordinate system have different units: in the vertical axis, we have a unit related to the instantaneous intensity of the wave, while in the horizontal axis we have the index at which each extremum occurs, ultimately a time-related unit.

@@

In this new coordinate system we have a basis vector $ \mathbf{b}^\prime = \{ \mathbf{b}^\prime_x, \mathbf{b}^\prime_y \}$ where $ \mathbf{b}^\prime_x = (i, 0) $ and $\mathbf{b}^\prime_y = (0, i) $, related to the original basis vector by equation \ref{eq:Basis}:

@@

The effect of this normalization can be seen in the points shown in figure \ref{fig:Vectors}, shown in scale in both coordinate systems. The set V of the vectors from each point of P to the next is also shown in the picture, and it will be used to estimate the average curvature of P. We define this set in equation \ref{eq:Vectors} below.

@@

Concerning direct methods \textcite{2014CarrollSurvey}, for example, derive three such definitions based on the approximation of a circle by an inscribed, centered and circumscribed polygon. In the context of three-dimensional meshes, \textcite{2016VasaMultivariate} evaluate a range of existing estimators from a multivariate point of view.

Those approaches define the discrete curvature in relation to the vertices of a wave (or mesh, in the three-dimensional case). In our particular case, it wouldn't make sense to talk about the curvature of a single pulse, since it is what our points ultimately represent. Instead, we are interested in the change of direction, what the term curvature ultimately means, between two adjacent pulses, as the vectors in the set V suggests.

@@

The rationale is to find, for each vector $ v_k \in \text{V} $, the radius $ r_k $ of the equivalent circle whose tangent has the same change in direction, in the same horizontal distance, as the vector of interest, as shown in figure \ref{fig:DiscreteCurvature}. The average curvature of P will then be obtained by the average of all radii $ r_k $.

From figure \ref{fig:DiscreteCurvature} is easy to see that $ r_k = v_{k,x} \sin(\theta_k)$ where $ v_{k,x} $ is the component of $ v_k $ in the horizontal direction; $ \theta_k $ itself can be obtained using the slope $ m_0 = 0 $ of a line in the horizontal direction and the slope $ m_k = v_{k,y} / v_{k,x} $ defined by $ v_k $ via the equality $ \theta_k = \arctan \big( (m_k - m_0) / (1 + m_k m_0) \big) $.

\begin{figure}[ht!]
  \centering
    \includegraphics{images/05DiscreteCurvature.pdf}
  \caption{The tangent unit vector of the circle changes from the horizontal direction in $ \hat{u}_0 $ to an inclination of $ \theta_k $ in $ \hat{u}_1 $, $ \theta_k $ being the angle that the vector $ v_k $ makes with the horizontal direction.}
  \label{fig:DiscreteCurvature}
\end{figure}

The radius of the circle that represents the average curvature of V (and P), and will be used to obtain the envelope of the discrete wave \textbf{w} can be obtained via equation \ref{eq:Radius}:

@@

We need now to construct an algorithm, using the radius just obtained, to identify the points that belong to the envelope. Generally, the algorithm for the alpha shapes approach resorts to the construction of the Delaunay triangulation of the set of points, that is latter filtered to contain only the outer edges. We can adopt a more straightforward method, as our points have a defined structure:

@@

For all circles of center $ x_o, y_o $ and radius $ r $ that can be constructed to pass through any two different points of P, that is, $ \forall \quad P_b, P_c $ such that $ (P_{b,x} - x_o)^2 + (P_{b,y} - y_o)^2 = (P_{c,x} - x_o)^2 + (P_{c,y} - y_o)^2 = r^2 $, $ P_a $ will be part of the set $ E $ of the points belonging to the envelope if and only if none of those circles contain $ P_a $, or $ P_a \in \text{E} \leftrightarrow (P_{b,x} - x_o)^2 + (P_{b,y} - y_o)^2 > r^2 $. Summarizing this, we have equation \ref{eq:InEnvelope}:

@@

The algorithm follows directly from the definition in \ref{eq:InEnvelope} after noting that the first and last members of P will always be part of the envelope, since they are reachable by a circle of any radius $ r $. From $ P_0 $, our first pivot point, we construct circles passing through $ P_1, P_2, \cdots, P_d $ until a circle that doesn't contain any point in P is found; $ P_d $ then becomes the new pivot point and is included in E, and this procedure continues until $ P_{m-1} $ is reached.

The procedure is formalized in the algorithm \ref{FindEnvelope}. We don't provide similar algorithmic descriptions of the preceding steps since the Python implementation can easily serve as pseudocode.

@@

The set E obtained via algorithm \ref{FindEnvelope} is a set of points. The most straightforward way to transform those points into the vector $ \textbf{e} \in \mathbb{R}^n $ is via linear interpolation, the approach used in this work. 
It is worth noting, however, that many other procedures are available, both for the interpolation and the smooth approximation of points.

\section{Results}
Given the lack of a precise definition of what an envelope should be, we begin this section suggesting an empirical metric, based on the behaviour of the carrier wave \textbf{c}. We also comment on some metrics presented in the literature, and compare our algorithm with traditional demodulation approaches, both in relation to execution time and accuracy; to this last end, a numerical indicator is also introduced.

@@

There are no agreed-upon objective metrics to assess the quality of an envelope. Recalling figure \ref{fig:Envelope}, however, and the accompanying discussion, an intuitive way is to consider the carrier wave \textbf{c}, and how well it seems to be bounded by the interval $ \{-1, 1\} $.

To illustrate how the algorithm performs on such metric, we provide the following figures of original discrete waves \textbf{w}, their extracted envelope \textbf{e}, and the inferred carrier, obtained after dividing, element-wise, the original wave by its envelope.

From figure \ref{fig:spoken_voice} it can be seen that the algorithm performs well under the carrier metric. This is especially remarkable for the case of spoken voice, a notorious complex wave with fast changes in pitch and high noise content.

These characteristics are in stark contrast with those of the sang voice in figure \ref{fig:singer_voice}, where a recording of an alto singer is shown; there, the wave is almost periodic, with a stable waveform.

@@

\begin{figure}[ht!]
  \centering
    \includegraphics{images/06Graph_spoken_voice.pdf}
  \caption{Signal, envelope and carrier for a record of male voice uttering the word “amazing”.}
  \label{fig:spoken_voice}
\end{figure}

@@

The first one states that, if a signal is bounded in magnitude, then its envelope should be, as well. Our algorithm complies with this requisite, as the envelope is composed of selected samples from the original signal itself.

@@

Nevertheless, if we picture both the wave and the envelope as samples from continuous functions, is easy to see that this is the case for our algorithm; since the envelope is composed by some samples of the digital wave, in the limit where all the samples of the wave are also part of the envelope, we have that the maximum frequency for both is the same.

@@

The third condition states what, for a periodic signal, the envelope should be a straight line with the intercept equal to the amplitude of this signal. Since all the local maxima have the same amplitude in the case of a periodic wave, our envelope will indeed be a straight line, with the same amplitude as the original wave. An example of such behaviour can be observed in figure \ref{fig:Comparison_sinusoid}.

@@

Direct comparison with many of the more recent algorithms is made difficult by the unavailability of digital implementations of such works, many designed to process analog signals \parencite[e.g.,][]{2018AssefModeling}.
